Cell Decompositions for Rank Two Quiver Grassmannians

### 2018-01-23 Grassmannian_cells.tex
Section 5 (not finished)

Recall that, given $\CC$-representations $X$ and $Y$ of $Q$, any tuple of linear maps $(g_\alpha:X_{s(\alpha)}\to Y_{t(\alpha)})_{\alpha\in Q_1}$ defines a short exact sequence $\ses{Y}{Z}{X}$ with middle term given by the vector spaces $Z_i=X_i\oplus Y_i$ for $i\in Q_0$ and the linear maps $Z_\alpha=\begin{pmatrix} X_\alpha & 0 \\ g_\alpha & Y_\alpha \end{pmatrix}$ for $\alpha\in Q_1$. 
In general, considering the linear map  
\[d_{X,Y}:\bigoplus_{i\in Q_0}\Hom_\CC(X_i,Y_i)\to\bigoplus_{\alpha\in Q_1}\Hom_\CC(X_{s(\alpha)},Y_{t(\alpha)}),\qquad (f_i)_{i\in Q_0}\mapsto(f_{t(\alpha)}\circ X_\alpha-Y_\alpha\circ f_{s(\alpha)})_{\alpha\in Q_1},\]
we have $\mathrm{ker}(d_{X,Y})=\Hom_Q(X,Y)$ and $\mathrm{coker}(d_{X,Y})=\Ext_Q(X,Y)$.

@@

The next step is to associate a $2$-quiver with certain (truncated) preprojective representation $ P_m^{\uj}=F_{K(n)}\tilde P_m^{\uj}$ for $\uj\in I_{n-1}$ whose strong successor closed subsets correspond to the cells of quiver Grassmannians. Therefore, we use the recursive construction described in Section \ref{Lifting}. In particular, we choose bases of the $\Ext$-groups which are compatible with the lifts to the universal covering. Recall that quiver Grassmannians $\Gr_{\tbfe}(P_m^V)$ attached to truncated preprojective representations only depend on $\tbfe$ and $\udim P_m^V$. In particular, in view of Theorem \ref{thm:truncpp}, these $2$-quivers can be seen as being attached to any (truncated) preprojective of this dimension.\sayT{Introduce notation for $2$-quivers associated with $P_m^V$?}.

For simplicity and in order to be in shape(?) with Sections \ref{Lifting} and \ref{sec:fibrations}, we work on the universal cover.
It turns out that all $2$-quivers can be built from the $2$-quivers of the lifts $\tilde P_1$, $\tilde P_2$ and $\tilde P_2^{[n-1]}=\tilde P_2^{(1,\ldots,n-1)}$ as defined in Section \ref{Lifting}, i.e. we have:
\[
  \tilde P_2^{[n-1]}=(1,\alpha_1)\xlongleftarrow{\alpha_1}(2,1).%=P_1^{(1)}=j^{(1)}_1\xlongleftarrow{v_1}i^{(1)}_1
\]
We define $P_m^{[k]}:=F_{K(n)}\tilde P_m^{[k]}$.
%In other words, $P_2^{(n-1)}$ is chosen as be the cokernel of a short exact sequence
%\[
%  \ses{ P_1^{n-1}}{ P_2}{P_2^{(n-1)}}.
%\]
Choosing the lifts of $P_1$ supported at $(1,\alpha_2),\ldots,(1,\alpha_n)$, we have 
\begin{align}\label{eq:case2}
  \Ext( P_2^{[n-1]},P_1)\cong \bigoplus_{i=2}^{n}\Ext(\tilde P_2^{[n-1]},\alpha_i\tilde P_{1})\cong\langle (2,1)\xrightarrow{\alpha_i}(1,\alpha_i)\mid i=2,\ldots,n\rangle.
\end{align}

This defines recursively the truncated preprojective representation $\tilde P_{2}^{[k]}$ for $k=n-1,\ldots,0$ with $\tilde P_2^{[0]}=\tilde P_2$ by 
\begin{align}
  \label{coeff}\xymatrix@R10pt@C20pt{&&(1,\alpha_1)\\ (2,1)\ar^{\alpha_1}[rru]\ar^{\alpha_{n-k}}[rrd] &&\vdots\\&&(1,\alpha_{n-k})}.
\end{align}
Note that, for general $m\geq 2$, we have a short exact sequence
\[\ses{\Hom(P_{m-1},P_{m-1})}{\Ext(P_m^{[k]},P_{m-1})}{\Ext(P_m^{[k+1]},P_{m-1})}.\]
Thus a basis of $\Ext(P_m^{[k]},P_{m-1})$ can be obtained by taking the last $k$ basis elements of a basis of $\Ext(P_m^{[n-1]},P_{m-1})$.
We define the $2$-quiver of $P_m^{[k]}$ by
\\

@@

\noindent
where the $2$-arrows are coloured red. This $2$-quiver is clearly equivalent to the quiver (\ref{coeff}). For clearness reasons, we will mostly work with the quiver instead.

The next step is to define the $2$-quivers of certain truncated preprojectives $\tilde P^{[k]}_3$ and $P^{[k]}_3$ whence the remaining $2$-quivers attached to truncated preprojectives $P_m^{[k]}$ can be obtained recursively.

@@

 This basis recursively gives rise to the following $2$-quiver of $\tilde P_3^{[k]}=\tilde P_3^{(1,\ldots,k)}$ (resp. for $P_3^{[k]}$) for $k=n-1,\ldots,0$ where we drop the indices of the vertices in abuse of notation. Note that the middle vertex of $\tilde P_3^{[n-1]}$ is $(1,1)$ and that all $2$-arrows end in this vertex of $\widetilde{K(n)}$.\\

>>>>

Recall that, given $\CC$-representations $X$ and $Y$ of $Q$, any tuple of linear maps $(g_\alpha:X_{s(\alpha)}\to Y_{t(\alpha)})_{\alpha\in Q_1}$ defines a short exact sequence $\ses{Y}{Z}{X}$ with middle term given by the vector spaces $Z_i=X_i\oplus Y_i$ for $i\in Q_0$ and the linear maps $Z_\alpha=\begin{pmatrix} X_\alpha & 0 \\ g_\alpha & Y_\alpha \end{pmatrix}$ for $\alpha\in Q_1$. 
In general, considering the linear map  
\begin{align}\label{eq:homext}d_{X,Y}:\bigoplus_{i\in Q_0}\Hom_\CC(X_i,Y_i)\to\bigoplus_{\alpha\in Q_1}\Hom_\CC(X_{s(\alpha)},Y_{t(\alpha)}),\qquad (f_i)_{i\in Q_0}\mapsto(f_{t(\alpha)}\circ X_\alpha-Y_\alpha\circ f_{s(\alpha)})_{\alpha\in Q_1},\end{align}
we have $\mathrm{ker}(d_{X,Y})=\Hom_Q(X,Y)$ and $\mathrm{coker}(d_{X,Y})=\Ext_Q(X,Y)$.

@@

The next step is to associate a $2$-quiver with certain lifts of (truncated) preprojective representation $\tilde P_m^{\uj}$ for $\uj\in I_{n-1}$ whose strong successor closed subsets correspond to the cells of quiver Grassmannians. Recall that the cells of the quiver Grassmannians attached to $P_m^{\uj}:= F_{K(n)}\tilde P_m^{\uj}$ coincide with those attached to the lift $\tilde P_m^{\uj}$.

For building these $2$-quivers, we use the recursive construction described in Section \ref{Lifting}. In particular, we choose bases of the $\Ext$-groups which are compatible with the lifts to the universal covering. Recall that quiver Grassmannians $\Gr_{\tbfe}(P_m^V)$ attached to truncated preprojective representations only depend on $\tbfe$ and $\udim P_m^V$. In particular, in view of Theorem \ref{thm:truncpp}, these $2$-quivers can be seen as being attached to any (truncated) preprojective of this dimension.

%For simplicity and in order to be in shape(?) with Sections \ref{Lifting} and \ref{sec:fibrations}, we work on the universal cover.
It turns out that all $2$-quivers can be built from the $2$-quivers of the lifts $\tilde P_1$, $\tilde P_2$ and $\tilde P_2^{[n-1]}:=\tilde P_2^{(1,\ldots,n-1)}$. Using the conventions of Section \ref{Lifting} the representation $\tilde P_2^{[n-1]}$ is defined by
\[
  \tilde P_2^{[n-1]}=(1,\alpha_1)\xlongleftarrow{\alpha_1}(2,1).%=P_1^{(1)}=j^{(1)}_1\xlongleftarrow{v_1}i^{(1)}_1
\]
We now recursively consider the lifted truncated preprojectives $\tilde P_m^{[k]}:=\tilde P_m^{(1,\ldots,k)}$ for $m\geq 2$ and $0\leq k\leq n-1$. We define $P_m^{[k]}:=F_{K(n)}\tilde P_m^{[k]}$.
%In other words, $P_2^{(n-1)}$ is chosen as be the cokernel of a short exact sequence
%\[
%  \ses{ P_1^{n-1}}{ P_2}{P_2^{(n-1)}}.
%\]
Choosing the lifts $\tilde P_{1,i}$ of $P_1$ supported at $(1,\alpha_i)$ for $2\leq i\leq n$ and using the description of the $\Ext$-groups induced by (\ref{eq:homext}), we have 
\begin{align}\label{eq:case2}
  \Ext( P_2^{[n-1]},P_1)\cong \bigoplus_{i=2}^{n}\Ext(\tilde P_2^{[n-1]},\tilde P_{1,i})\cong\langle (2,1)\xrightarrow{\alpha_i}(1,\alpha_i)\mid i=2,\ldots,n\rangle.
\end{align}

This yields recursively the truncated preprojective representation $\tilde P_{2}^{[k]}$ for $k=n-1,\ldots,0$ with $\tilde P_2^{[0]}=\tilde P_2$ by 
\begin{align}
  \label{coeff}\xymatrix@R10pt@C20pt{&&(1,\alpha_1)\\ (2,1)\ar^{\alpha_1}[rru]\ar^{\alpha_{n-k}}[rrd] &&\vdots\\&&(1,\alpha_{n-k})}.
\end{align}
For general $m\geq 2$, there exists a short exact sequence
\[\ses{\Hom(P_{m-1},P_{m-1})}{\Ext(P_m^{[k]},P_{m-1})}{\Ext(P_m^{[k+1]},P_{m-1})}.\]
Thus a basis of $\Ext(P_m^{[k]},P_{m-1})$ can be obtained by taking the last $k$ basis elements of a basis of $\Ext(P_m^{[n-1]},P_{m-1})$.
We define the $2$-quiver of $P_2^{[k]}$ by
\\

@@

\noindent
where the $2$-arrows are coloured red. This $2$-quiver is clearly equivalent to the quiver (\ref{coeff}). For clearness reasons, we will mostly work with the quiver (\ref{coeff}) instead.

The next step is to define the $2$-quivers of certain truncated preprojective lifts $\tilde P^{[k]}_3$ (resp. of $P^{[k]}_3$) whence the remaining $2$-quivers attached to the truncated preprojectives $P_m^{[k]}$ can be obtained recursively.

@@

 This basis recursively gives rise to the following $2$-quiver of $\tilde P_3^{[k]}=\tilde P_3^{(1,\ldots,k)}$ (resp. for $P_3^{[k]}$) for $k=n-1,\ldots,0$ where we drop the indices of the vertices in abuse of notation. Note that the middle vertex of $\tilde P_3^{[n-1]}$ is $(2,\alpha_n^{-1})$ and that all $2$-arrows end in the vertex $(1,1)$ of $\widetilde{K(n)}$.\\
